\bf Hierarchical Network Structure Promotes Dynamical Robustness

### 2015-05-21 tex/figures.tex
add new robustnessconcept fig
update: incorporate references to new conceptual robustness figures

\begin{figure*}[!ht]
\centering
\noindent\includegraphics[width=0.8\textwidth]{fig/biomodelexamples.pdf}
\caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields. The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here. The third row shows the general form of a system of differential equations that can be used to model any network architecture within each domain.}
\label{fig:biomodelexamples}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\noindent\includegraphics[width=0.8\textwidth]{fig/modsccsym.pdf}
\caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems. (B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems. The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them. We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one. The two panels in the middle represent examples of hierarchical modular systems that posess both modularity (i.e. SCCs with more than one variable) and hierarchy. (C) Symmetries of the $\hier$ transformation between graphs and SCCs. The transformation $a$ represents an interchange of SCCs, $b$ moving a link between nodes in a component and $c$ adding a link. All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
\label{fig:modsccsym}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\noindent\includegraphics[width=0.8\textwidth]{fig/combinedfigs.pdf}
\caption{{\bf Characterization of stability and robustness according to properties of system structure for three variable systems} (A) Robustness versus connectivity. The red line represents a best fit in the least-squares sense with Pearson product-moment correlation coefficient $r=0.29$. The lowest and highest robustness network architectures are labelled. Other network architectures are shown in \ref{tab:structstabmat3}. (B) Robustness versus hierarchy. Correlation coefficient $r=0.67$. (C) Number of cycles and (D) hierarchy vs connectivity and robustness. The color of each point represents the average robustness of all graphs having the parameters specified on the $x$ and $y$ axes.
}
\label{fig:combined}
\end{figure*}

>>>>

\begin{figure*}[!ht]
\centering
\noindent\includegraphics[width=1.0\textwidth]{fig/biomodelexamples.pdf}
\caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields. The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here. The third row shows the general form of a system of differential equations that can be used to model any network architecture within each domain.}
\label{fig:biomodelexamples}
\end{figure*}

@@

\begin{figure*}[!htbp]
\centering
\noindent\includegraphics[width=1.0\textwidth]{fig/robustnessconceptvis.pdf}
\caption{{\bf Stability and equivalence classes of fixed points.} Linearization around a steady-state of a model of a biological network allows for the assessment of the stability of that state. For two-dimensional systems, assessment of stability in terms of the trace and determinant of the Jacobian matrix associated to a particular steady-state geometrically partitions the plane into a stable region (S, blue) and an unstable region (U, red). For larger systems in higher dimensions this geometry is too complicated to visualize, but it is nevertheless well-defined and can be investigated algebraically and numerically. (top center) Jacobian matrices such as $A_1$ partition the space of fixed points into equivalence classes $[A_1]$, each of which is either stable or unstable.}
\label{fig:robustnessconcept}
\end{figure*}

@@

\begin{figure*}[!htbp]
\centering
\noindent\includegraphics[width=1.0\textwidth]{fig/robustnessprocess.pdf}
\caption{{\bf Assessment of dynamical robustness.} (A) Any stochastic process induced on the stable and unstable regions of the space of fixed points can be represented as a state transition diagram. The probability of remaining in the stable region in the context of such modifications, given a previously existing stable state, is then provided by estimating $R$. (B) The stochastic process that results from modifications to a given dynamical model over evolutionary timescales induces a stochastic process on the equivalence classes of fixed points $\mathcal{F}/{\sim}$ represented by the boxes in the top row. This in turn induces a process on the stable and unstable regions, which can be used to estimate robustness. (C) Each network architecture encoded by a directed graph, $\{G_1, \ldots, G_k \}$, that indicates the manner in which the variables of a biological network depend upon one another selects a particular subset of equivalence classes from $\mathcal{F}/{\sim}$ (see \ref{eq:jacgrapheqs}). Robustness can therefore be estimated for each network architecture independently.}
\label{fig:robustnessprocess}
\end{figure*}

\begin{figure*}[!ht]
\centering
\noindent\includegraphics[width=1.0\textwidth]{fig/modsccsym.pdf}
\caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems. (B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems. The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them. We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one. The two panels in the middle represent examples of hierarchical modular systems that posess both modularity (i.e. SCCs with more than one variable) and hierarchy. (C) Symmetries of the $\hier$ transformation between graphs and SCCs. The transformation $a$ represents an interchange of SCCs, $b$ moving a link between nodes in a component and $c$ adding a link. All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
\label{fig:modsccsym}
\end{figure*}

\begin{figure*}[!ht]
\centering
\noindent\includegraphics[width=1.0\textwidth]{fig/combinedfigs.pdf}
\caption{{\bf Characterization of stability and robustness according to properties of system structure for three variable systems} (A) Robustness versus connectivity. The red line represents a best fit in the least-squares sense with Pearson product-moment correlation coefficient $r=0.29$. The lowest and highest robustness network architectures are labelled. Other network architectures are shown in \ref{tab:structstabmat3}. (B) Robustness versus hierarchy. Correlation coefficient $r=0.67$. (C) Number of cycles and (D) hierarchy vs connectivity and robustness. The color of each point represents the average robustness of all graphs having the parameters specified on the $x$ and $y$ axes.
}
\label{fig:combined}
\end{figure*}

### 2015-05-21 tex/introduction.tex
update: incorporate references to new conceptual robustness figures

%!TEX root = ../paper.tex
The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to derive a system of differential equations to model a particular biological network, attempt to fit that model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}. All of these models utilize essentially equivalent mathematical structures, \ref{fig:biomodelexamples}, \cite{RossCr2003,Palsson2011a,Sauro2012}. Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
% \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}

@@

Here we determine the relationship between network hierarchy, a topological property, and \emph{robustness}, a dynamical property. Robustness is of interest in biological systems at all scales, and has been previously studied in the context of biochemical networks \cite{Alon1999,Shinar2010}, gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Draghi2010,Wagner2013}, and ecological networks \cite{Rohr2014}. Over physiological timescales, the robustness of a particular network state may be evaluated by determining its linear stability \cite{Davis1962}. Network states that are linearly stable, are robust to perturbations in the states of their components. For example, in the case of gene-regulatory networks, a state vector containing protein molecule counts or concentrations of each gene that is stable with respect to the dynamics of the network will exponentially suppress any relatively small modifications to the state of a gene and return to the initial stable state. Here we are interested in evaluating robustness over evolutionary as opposed to physiological timescales.

Over evolutionary timescales it is expected for there to be fluctuations, not only in the state, but in any aspect of the model specifying the dynamics of the network itself (i.e. parameters, structure of the rate functions, etc.). It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states. However, what must remain invariant on such evolutionary timescales is the higher-level constraint that the system possess a relatively high overall probability of remaining in a stable state upon modifications to the underlying dynamical process. This is necessary in order for networks of lower-level components to exist, regardless of what state they exist in, long enough to serve as the substrate out of which networks of relatively higher-level components are constructed \cite{Simon2002}. What is important over at least moderate evolutionary timescales is then the conditional probability, $R$, given a system is in a stable state, that upon modifications to its structure in the context of environmental fluctuations, it remains stable, regardless of where the stable state is located within the state space \ref{fig:robustnessconcept}. For the purpose of this investigation then, we quantify dynamical robustness in this way.
% as the probability of stability to potentially large perturbations of the model specifying the system dynamics for systems which have already been determined to be linearly stable.

>>>>

%!TEX root = ../paper.tex
The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to derive a system of differential equations to model a particular biological network, attempt to fit that model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}. All of these models utilize essentially equivalent mathematical structures, \ref{fig:biomodelexamples}, \cite{RossCr2003,Palsson2011a,Sauro2012}, and developing unified mathematical descriptions of them is one of the paramount goals of systems biology.
% \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}

@@

Here we determine the relationship between network hierarchy, a topological property, and \emph{robustness}, a dynamical property. Robustness is of interest in biological systems at all scales, and has been previously studied in the context of biochemical networks \cite{Alon1999,Shinar2010}, gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Draghi2010,Wagner2013}, and ecological networks \cite{Rohr2014}. Over physiological timescales, the robustness of a particular network state may be evaluated by determining its linear stability \cite{Davis1962}. Network states that are linearly stable, are robust to perturbations in the states of their components. For example, in the case of gene-regulatory networks, a state vector containing protein molecule counts or concentrations of each gene that is stable with respect to the dynamics of the network will exponentially suppress any relatively small modifications to the state of a gene and return to the initial stable state.

Here, rather than evaluating system stability for a particular model of a biological network over physiological timescales, we are interested in evaluating robustness over evolutionary timescales where the form of the most accurate underlying model is itself subject to change. Over evolutionary timescales it is expected for there to be fluctuations, not only in the state, but in any aspect of the model specifying the dynamics of the network itself (i.e. parameters, structure of the rate functions, etc.). It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states. However, what must remain invariant on such evolutionary timescales is the higher-level property that the system possess a relatively high overall probability of remaining in a stable state upon modifications to the underlying dynamical process subject to environmental constraints. This is necessary in order for networks of lower-level components to exist, regardless of what state they exist in, long enough to serve as the substrate out of which networks of relatively higher-level components are constructed \cite{Simon2002}. What is important over at least moderate evolutionary timescales is then the conditional probability, $R$, given a system is in a stable state, that upon modifications to its structure in the context of environmental fluctuations, it remains stable, regardless of where the stable state is located within the state space \ref{fig:robustnessconcept}. For the purpose of this investigation then, we quantify dynamical robustness in this way.
% as the probability of stability to potentially large perturbations of the model specifying the system dynamics for systems which have already been determined to be linearly stable.

### 2015-05-21 tex/overview.tex
update: incorporate references to new conceptual robustness figures

In the general case of a dynamical system with $n$ components, where the components may be concentrations of chemical species, genes, or biological species, we have an $n$-dimensional vector of state variables or observables $(x_1(t), \ldots x_n(t)) = \vec{x}(t)
$
whose components are solutions to the arbitrary first order system
\begin{equation}\label{eq:eom}
\frac{dx_i(t)}{dt} = F_i(\vec{x}(t), \vec{p}), \; (i=1,\ldots,n)
\end{equation}
where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p} \in \mathbb{R}^s$ is the vector of $s$ parameters of the $F_i$. These parameters typically represent reaction rates in chemical and gene-regulatory networks, interaction strengths in ecological networks, or cooperativity exponents. For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (n_1, \ldots, n_N)$, $\vec{p}=(r_1,\ldots,r_N,b_{11},\ldots,b_{NN})$, and $F_i = r_i n_i + \sum_{j=1}^{N} b_{ij} n_i n_j$. The set of all such dynamical systems, $\mathcal{D}$, is then
\begin{equation}\label{eq:setdynsys}
\mathcal{D} = \{ (F,\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
\end{equation}

Fixed points are the simplest class of solutions to the dynamical system characterizing its long-term behavior. If $\vec x$ is a fixed point (i.e. $F_i(\vec{x})=0$ for all $i$), we
may proceed to ask whether it is dynamically stable.
Intuitively, dynamic stability means that, if one chooses the initial
conditions sufficiently close to the fixed point, the solution will
stay nearby.  Physically, this is important because,
if a fixed point ${\vec x}^0$ is unstable, we have zero probability of
observing the solution ${\vec x}(t) = {\vec x}^0$ in the absence of coupling to another system. The Lotka-Volterra model has two fixed points: the trivial one of all zero species $n_i=0$ and the other given implicitly by $r_i + \sum_{j=1}^{N} b_{ij} n_j = 0$. The set of all such fixed points, $\mathcal{F}$, is then
\begin{equation}\label{eq:setdynsys}
\mathcal{F} = \{ (F,\vec{p},\vec{x}^0) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s, F(\vec{x}^0,\vec{p})=0 \}.
\end{equation}

@@

Jacobian matrices define an equivalence relation on fixed points given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}',\vec{x}^{0'}) $ if and only if
\begin{equation}\label{eq:jaceqrel}
\left. \frac{\partial F(\vec{x},\vec{p})}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^0} =
\left. \frac{\partial F'(\vec{x},\vec{p}')}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^{0'}}.
\end{equation}
This relation then partitions the set of all fixed points into equivalence classes where the class $[A]$ of $\mathcal{F} / \sim$ associated to Jacobian matrix $A \in \mathbb{R}^{n \times n}$ is
\begin{equation}\label{eq:jaceqs}
[A] = \left\{ (F,\vec{p},\vec{x}^0) \; | \; \left. \frac{\partial F(\vec{x},\vec{p})}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^0} = A \right\}.
\end{equation}

% \subsection{Overview}
The dynamical model given in terms of a system of differential equations for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
For a general system $X \in \mathcal{D}$ the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ if $F_i \hbox{ depends on } x_j$ and $0$ if $F_i \hbox{ does not depend on } x_j$. These two conditions on global system interactions are expressed respectively for all $\vec{x}$ in terms of elements of the Jacobian matrix as $\frac{\partial F_i}{\partial x_j}(\vec{x}) \neq 0$ and $\frac{\partial F_i}{\partial x_j}(\vec{x}) = 0$.
%  \begin{displaymath}
%    \adj(G_X)_{ij} = \left\{
%      \begin{array}{ll}
%        1, & F_i \hbox{ depends on } x_j\\
%        0, & F_i \hbox{ does not depend on } x_j
%      \end{array}.
%    \right.
% \end{displaymath}
For large systems, since any given component is only likely to interact with a relatively small proportion of the other components, these matrices may be sparse.
We can also associate a local interaction graph $G_A$ given by an adjacency matrix $\adj(G_A)$ to each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ where
 \begin{equation}\label{eq:lininterdepadj}
   \adj(G_A)_{ij} = \left\{
     \begin{array}{lr}
       1, & a_{ij} \neq 0\\
       0, & a_{ij} = 0
     \end{array}.
   \right.
\end{equation}
In general, the graph $G_A$ is a subgraph of $G_X$, however, $G_A$ is almost always equivalent to $G_X$ (\refsupp{}). We define the connectivity to be equal to the number of edges in $G_A$, which is equivalent to summing up the number of non-zero entries of $\adj(G_A)$.
These interaction graphs can be viewed as deriving from the combination of system components that accept a given pattern of inputs and produce a given pattern of outputs (\reffigexamplesystemmodules).

Each distinct directed graph $G$ selects a subset of $\mathcal{F} / \sim$
\begin{equation}\label{eq:jaceqs}
[G] = \left\{ [A] \; | \; G_A = G \right\}.
\end{equation}
The $G$-classes thereby partition the collection of fixed points, $\mathcal{F}$, over the space of dynamical systems, $\mathcal{D}$, according to the interactions among the variables of the dynamical system represented by the topology of $G$. This partition, $\mathcal{F} / G$, is a coarsening of $\mathcal{F} / \sim$.

@@

In the process of biological evolution, the parameter values $\vec{p}$ or the form of the functions $F_i$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due, for example, to mutations or environmental fluctuations. These changes may alter the capacity of a given system to achieve a stable steady state. In order to evaluate this we consider a stochastic process whereby for a given dynamical system associated to a stable matrix $A$, we consider a perturbation of the system leading to $A'$ where $\adj(G_A) = \adj(G_{A'})$ \ref{fig:robustnessconcept}. This corresponds to a dynamical system whose connectivity is preserved, but where the underlying model is defined in terms of a different vector of parameters ${\vec{p}}\,'$ (\refsupp{}). We then ask, for each distinct network topology on $n$ components, what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix. This probability is what we refer to as dynamical robustness.

We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$. The entries of the Jacobian are sampled independently from a generic probability distribution. Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:jacobianvectorfields}. The expression in \ref{eq:sccrobustness} holds for the case of simultaneously perturbing any number of elements of the Jacobian, but can be schematized for the case of perturbing a single element as \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)
\begin{equation}\label{eq:robschematic}
R_{tot} = \frac{l+d_1 R_1 + d_2 R_2 + \cdots}{l+d_1 + d_2 + \cdots}.
\end{equation}
For instance, if our graph is the one in \reffigscc{} (middle panels), then we have two connected components, one with two nodes, and one with one node.  From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and robustness $0.62$.  The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.  Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is computed from \ref{eq:robschematic} in \reffigscc{}, which agrees with the value computed in \ref{tab:structstabmat} up to sampling error.

>>>>

In the general case of a dynamical system with $n$ components, where the components may be concentrations of chemical species, genes, or biological species, we have an $n$-dimensional vector of state variables or observables $(x_1(t), \ldots x_n(t)) = \vec{x}(t)
$
whose components are solutions to the arbitrary first order system
\begin{equation}\label{eq:eom}
\frac{dx_i(t)}{dt} = F_i(\vec{x}(t), \vec{p}), \; (i=1,\ldots,n)
\end{equation}
where $F=\{F_1,\ldots,F_i,\ldots,F_n\}$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p} \in \mathbb{R}^s$ is the vector of $s$ parameters of the $F_i$. These parameters typically represent reaction rates in chemical and gene-regulatory networks, interaction strengths in ecological networks. For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (n_1, \ldots, n_N)$, $\vec{p}=(r_1,\ldots,r_N,b_{11},\ldots,b_{NN})$, and $F_i = r_i n_i + \sum_{j=1}^{N} b_{ij} n_i n_j$. The set of all such dynamical systems, $\mathcal{D}$, for a given number of state variables $n$ and a given number of parameters $s$, is then
\begin{equation}\label{eq:setdynsys}
\mathcal{D} = \{ (F,\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
\end{equation}

Fixed points are the simplest class of solutions to the dynamical system characterizing its long-term behavior. If $\vec x$ is a fixed point (i.e. $F_i(\vec{x})=0$ for all $i$), we
may proceed to ask whether it is dynamically stable.
Intuitively, dynamic stability means that, if one chooses the initial
conditions sufficiently close to the fixed point, the solution will
stay nearby.  Physically, this is important because,
if a fixed point ${\vec x}^0$ is unstable, we have zero probability of
observing the solution ${\vec x}(t) = {\vec x}^0$ in the absence of coupling to another system. The Lotka-Volterra model has two fixed points: the trivial one of all zero species $n_i=0$ and the other given implicitly by $r_i + \sum_{j=1}^{N} b_{ij} n_j = 0$. The set of all such fixed points, $\mathcal{F}$, is then
\begin{equation}\label{eq:setfixedpoints}
\mathcal{F} = \{ (F,\vec{p},\vec{x}^0) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s, F(\vec{x}^0,\vec{p})=0 \}.
\end{equation}

@@

Jacobian matrices define an equivalence relation, $\sim$, on fixed points given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}',\vec{x}^{0'}) $ if and only if
\begin{equation}\label{eq:jaceqrel}
\left. \frac{\partial F(\vec{x},\vec{p})}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^0} =
\left. \frac{\partial F'(\vec{x},\vec{p}')}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^{0'}}.
\end{equation}
This relation then partitions the set of all fixed points into equivalence classes $\mathcal{F}/{\sim}$, where the class $[A]$ associated to Jacobian matrix $A \in \mathbb{R}^{n \times n}$ is
\begin{equation}\label{eq:jaceqs}
[A] = \left\{ (F,\vec{p},\vec{x}^0) \; | \; \left. \frac{\partial F(\vec{x},\vec{p})}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^0} = A \right\}.
\end{equation}
An example with two members of $\mathcal{F}$ from each of four different equivalence classes $[A_1]$, $[A_2]$, $[A_3]$, and $[A_4]$ of $\mathcal{F}/{\sim}$ is shown in \ref{fig:robustnessconcept}.

% \subsection{Overview}
The interactions among variables in a dynamical model for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
For a general system $X \in \mathcal{D}$ the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ if $F_i \hbox{ depends on } x_j$ and $0$ if $F_i \hbox{ does not depend on } x_j$. These two conditions on global system interactions are expressed respectively for all $\vec{x}$ in terms of elements of the Jacobian matrix as $\frac{\partial F_i}{\partial x_j}(\vec{x}) \neq 0$ and $\frac{\partial F_i}{\partial x_j}(\vec{x}) = 0$.
%  \begin{displaymath}
%    \adj(G_X)_{ij} = \left\{
%      \begin{array}{ll}
%        1, & F_i \hbox{ depends on } x_j\\
%        0, & F_i \hbox{ does not depend on } x_j
%      \end{array}.
%    \right.
% \end{displaymath}
For large systems, since any given component is only likely to interact with a relatively small proportion of the other components, these matrices may be sparse.
We can also associate a local interaction graph $G_A$ given by an adjacency matrix $\adj(G_A)$ to each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ where
 \begin{equation}\label{eq:lininterdepadj}
   \adj(G_A)_{ij} = \left\{
     \begin{array}{lr}
       1, & a_{ij} \neq 0\\
       0, & a_{ij} = 0
     \end{array}.
   \right.
\end{equation}
In general, the graph $G_A$ is a subgraph of $G_X$, however, $G_A$ is almost always equivalent to $G_X$ (\refsupp{}). We define the connectivity to be equal to the number of edges in $G_A$, which is equivalent to summing up the number of non-zero entries of $\adj(G_A)$.
These interaction graphs can be viewed as deriving from the combination of system components that accept a given pattern of inputs and produce a given pattern of outputs (\reffigexamplesystemmodules).

Each distinct directed graph $G$, of which there are $k=2^{n(n-1)}$ that could be associated to the interactions in a model defined on $n$ variables, selects a subset of $\mathcal{F} / {\sim}$ (see~\ref{fig:robustnessprocess}C)
\begin{equation}\label{eq:jacgrapheqs}
[G] = \left\{ [A] \; | \; G_A = G \right\}.
\end{equation}
The $G$-classes thereby partition the collection of fixed points, $\mathcal{F}$, over the space of dynamical systems, $\mathcal{D}$, according to the interactions among the variables of the dynamical system represented by the topology of $G$. This partition, $\mathcal{F} / G$, is a coarsening of $\mathcal{F} / {\sim}$.

In the process of biological evolution, the parameter values $\vec{p}$, form of the functions $F$, and environmental conditions restricting access to different fixed points $\vec{x}^0$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due to environmental fluctuations. These modifications induce a stochastic process on the equivalence classes of fixed points $\mathcal{F}/{\sim}$ \ref{fig:robustnessprocess}B. These changes may likewise alter the capacity of a given system to achieve a stable steady state, thereby inducing a coarse-grained stochastic process on the stable regions of the space of fixed points \ref{fig:robustnessprocess}A and B. In order to evaluate this we consider a process whereby perturbations are applied to a given dynamical system and fixed point associated to a stable matrix $A$ leading to $A'$. In order to classify the properties of this process according to network architecture, we enforce the condition $\adj(G_A) = \adj(G_{A'})$, which results in analogous processes defined on each $G$-class of $\mathcal{F}/G$ \ref{fig:robustnessconcept}C. This corresponds to an ensemble of dynamical systems with equivalent connectivities, but where each model in the ensemble may otherwise be defined in terms of a different collection of rate functions $F'$, vector of parameters ${\vec{p}}\,'$, or environmental conditions restricting access to $\vec{x}^{0'}$ (\refsupp{}). We then ask, for each distinct network topology on $n$ components, what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix. This conditional probability corresponds to the parameter $R$ of the two-state process depicted in \ref{fig:robustnessprocess}A, and it is what we refer to as dynamical robustness. Comparing the values of $R$ for each $G$-class in $\mathcal{F}/G$ places a partial ordering on network architectures, $G$, which allows for the determination of which network architectures would be expected to be enriched in a process selecting for higher $R$ values.

@@

We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$. The entries of the Jacobian are sampled independently from a generic probability distribution. Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}. The expression in \ref{eq:sccrobustness} holds for the case of simultaneously perturbing any number of elements of the Jacobian, but can be schematized for the case of perturbing a single element as \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)
\begin{equation}\label{eq:robschematic}
R_{tot} = \frac{l+d_1 R_1 + d_2 R_2 + \cdots}{l+d_1 + d_2 + \cdots}.
\end{equation}
For instance, if our graph is the one in \reffigscc{} (middle panels), then we have two connected components, one with two nodes, and one with one node.  From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and robustness $0.62$.  The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.  Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is computed from \ref{eq:robschematic} in \reffigscc{}, which agrees with the value computed in \ref{tab:structstabmat} up to sampling error.
